NumPy---Array Computation for Python

### 2020-01-10 body.tex
Work on array outline
Another pass on NumPy array

The NumPy array is a structure that efficiently stores and accesses
$n$-dimensional array data\cite{vanderwalt2011numpy}.
The need for such a data structure arises often in science.
For instance, it may be used to store experimental measurements made at
discrete time intervals (1-D arrays), gray-level photographs (2-D arrays),
spectral measurements on a regular grid (3-D arrays), or color videos (4-D
arrays).

The NumPy array consists of a pointer to memory, along with meta-data used to
interpret the data stored there.
Computer memory stores elements linearly and has no notion
of higher-dimensions, but we can interpret it as an $n$-dimensional array by also
providing the array's shape, strides (the number of bytes to advance in memory to skip
along rows, columns, etc.), and data type (see Figure~\ref{fig:array-concepts}).

Data types (dtypes for short) tell us more about the elements stored in a NumPy
array.
% can hold ...?, e.g.,
Arrays can hold, e.g., numerical data (of lower and higher precision), strings,
timestamps, or Python objects, each identified by its own dtype.

@@

Users predominantly interact with NumPy arrays using {\em indexing} (to access sub-arrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$, and $@$, for matrix multiplication), together with {\em utility functions}; together, these provide an easily readable, expressive, high-level API (Application Programming Interface) for array manipulation.

>>>>

The NumPy array is a structure that efficiently stores and accesses
$n$-dimensional arrays, also known as tensors\cite{vanderwalt2011numpy}.
The need for such a data structure arises often in science, as it can be used to represent everything from time-series measurements to time-varying spectral measurements on a grid.

The NumPy array consists of a pointer to memory, along with meta-data used to
interpret the data stored there, notably {\em data type}, {\em shape}, and {\em strides} (see Figure~\ref{fig:array-concepts}).

A {\em data type} describes the nature of elements stored in an array.  An array has a single data type, and each array element occupies the same number of bytes in memory.  Examples of data types include real and complex numbers (of lower and higher precision), strings, timestamps, and pointers to Python objects.

The {\em shape} of an array determines the number of axes, or dimensionality, of the array.  For example, a vector of numbers can be stored as a 1-D array of shape $N$, while color videos are 4-D arrays of shape $T \times M \times N \times 3$.

{\em Strides} are necessary to interpret computer memory, which stores elements linearly, as $n$-dimensional arrays.  It describes the number of bytes to move forward in memory to jump from row to row, column to column, and so forth.  Consider, for example, a 2-D array of floating point numbers with shape $4 \times 3$, where each element occupies 8 bytes in memory.  To move between consecutive columns we need to jump forward 8 bytes in memory, and to access the next row $3 \times 8 = 24$ bytes.  The strides of that array are therefore $(24, 8)$.

Portions of an array can be accessed through an {\em indexing} operation.  Arrays can be indexed in several ways to return single elements, sub-arrays, or elements that satisfy a specific condition.  Arrays can even be indexed using other arrays (see Figure~\ref{fig:array-concepts}).  Wherever possible, indexing that retrieves a subarray returns a {\em view} on the original array, such that data is shared between the two arrays.  This provides a powerful way to operate on subsets of array data, while limiting memory usage.

@@

Users predominantly interact with NumPy arrays using {\em indexing} (to access sub-arrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$, and, for matrix multiplication, $@$), together with {\em utility functions}; together, these provide an easily readable, expressive, high-level API (Application Programming Interface) for array manipulation.
